---
title: "Homework 4"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(deSolve)
library(patchwork)
```

```{r}
knitr::opts_chunk$set(
        echo = TRUE,
        warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() +theme(legend.position = "bottom"))

options(
  ggplot2.continuous.color = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_color_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r}
SIRdem=function(t,state,parameters){
  with(as.list(c(state,parameters)),{
    
    # rate of change
    dS=mu*N-beta*S*I/N-mu*S
    dI=beta*S*I/N-gamma*I-mu*I
    
    # return the rate of change
    list(c(dS,dI))
  }) # end with(as.list...)
}
```


# **Part 1: SIR Model with Demography**
The SIR model with demography provides a way to study endemic diseases, in particular what proportion of the population are susceptible or infected when the disease becomes endemic. We've provided the code for both the basic SIR model with demogrpahy in the script. Do the following simulations and report your findings.

Run both models for 100 years, using the following initial conditions:
N = 1e5
I0 = 20
S0 = 0.2 * N
beta = 520 / 365
gamma = 1 / 7
mu = 1 / 70 / 365

```{r}
N=1e5
I0=20
S0=0.2*N
state=c(S=S0,I=I0)
parameters = c(beta=520/365, gamma=1/7, mu=1/(70*365))
times=seq(0,365*100,by=1)
```

```{r}
percent_sim = ode(y = state, times = times, func = SIRdem, parms = parameters)
percent_sim_df = as.data.frame(percent_sim)

percent_sim_df$S_frac = percent_sim_df$S / N
percent_sim_df$I_frac = percent_sim_df$I / N
```

## Question 1
Plot *%I* vs *time* for both models and compare. What is the main difference and why?

```{r}
if(F){

  parameters.SIR =  function(t, state, parameters){
    with(as.list(c(state, parameters)),{
      
      #rate of change
      dS = -beta * S * I / N
      dI = beta * S * I / N - gamma * I
      
      #return the rate of change
      list(c(dS, dI))
    })
  }
  
  parameters.SIRdem =  function(t, state, parameters){
    with(as.list(c(state, parameters)),{
      
      #rate of change
      dS = mu * N - beta * S * I / N - mu * S
      dI = beta * S * I/N - gamma * I - mu * I
      
      #return the rate of change
      list(c(dS, dI))
    })
}}
```

```{r}
percent_infectious = ggplot(percent_sim_df, aes(x = time, y = I_frac, color = time)) +
  geom_line() +
  labs(title = "Fraction of Infectious Population", 
       x = "Day",
       y = "Number of People")

percent_infectious
```


## Question 2
What are the %S (S/N), and %I (I/N) at the end of the simulation? How are these values (i.e. S/N, and I/N at equilibrium) compared to those computed by the formula we derived in the class?
Hint: what is R0 for the SIR model with demography?

```{r}
percent_susceptible = ggplot(percent_sim_df, aes(x = time, y = S_frac, color = time)) +
  geom_line() +
  labs(title = "Fraction of Susceptible Population", 
       x = "Day",
       y = "Number of People")

percent_susceptible
```

```{r}
percent_infectious + percent_susceptible
```

# **Part 2: SIS Model**

## Question 3
Complete the SIS model [include your code for the SIS model in your report]

```{r}
SIS=function(t,state,parameters){
  with(as.list(c(state,parameters)),{
    # rate of change
    
    # EQUATIONS for the SIS:
    dS=-beta*S*I/N+gamma*I
    dI= beta*S*I/N-gamma*I
    
    # return the rate of change
    list(c(dS,dI))
  }) # end with(as.list...)
}
```

```{r}
N=1e5
I0=10
S0=N-I0
state=c(S=S0,I=I0)
parameters=c(beta=.5,gamma=.3)

times=seq(0,100,by=1)
sim=ode(y=state,times=times,func=SIS,parms=parameters)
s=sim[,'S']/N
i=sim[,'I']/N
```


## Question 4
Run the SIS model using the following initial conditions:
N = 1e5
I0 = 10
S0 = N-I0
beta = 0.5
gamma = 0.3
times = 0:100

Plot *S* vs time and *I* vs time

```{r}
SIS = ode(y = state, times = times, func = SIS, parms = parameters)
SIS_df = as.data.frame(SIS)
```

```{r}
sis_sus = ggplot(SIS_df, aes(x = time, y = S, color = time)) +
  geom_line() +
  labs(title = "Susceptible Population", 
       x = "Day",
       y = "Number of People") 

sis_sus
```

## Question 5
What are S/N and I/N at equilibrium? How are these values compared to results from the formula derived from the mdel (I*/N = 1-1/R0 and S*/N = 1/R0)?
Hint: what is R0 for the SIS model?


# **Part 3: SEIR Model**

## Question 6
Code up an SEIR model based on the equations above [provide your code for the SEIR model in your report].

## Quetsion 7
Compare S, I, Incidence, over time computed by the 2 models. What are the differences between the two simulations?


# **Part 4: Application: Ebola Epidemic**



